|

### 2015-12-08 manuscript/manuscript_pve.Rmd
added chapter heading
added chapter#s to section

removed key words and author correspondence

---
title: "Below-ground sink limitation alters growth and carbon balance of Eucalyptus seedlings"
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle_thesis.docx
  pdf_document: default
  html_document:
    number_sections: yes

@@

\
*Corresponding author*: 
Courtney Campany
E: courtneycampany@gmail.com

@@

# Abstract {.unnumbered}

@@

## Keywords {.unnumbered}
photosynthesis, growth, sink regulation, carbon allocation, soil volume

# Introduction

@@

# Materials and Methods

## Experimental design
This experiment was located at the Hawkesbury Forest Experiment site in Richmond, NSW, Australia. Plots were located in an open cover paddock that was converted from native pasture grasses. Top soils at this site are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg^-1^ total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity. At this site a soil hard layer exists at ~1.0 m with a transition to heavy clay soils. The climate for the region is classified as sub-humid temperate.

@@

## Growth and morphology metrics
Seedlings were planted in summer (January 21^st^ 2013) and stem height, diameter at 15 cm and leaf count were measured weekly thereafter. Once the growth rate of individual plants had significantly declined a full biomass harvest was completed and the experiment ended (May 21^st^ 2013). Dry mass of leaves, stems, roots and total leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) were measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing total measured leaf area by total leaf count of only fully expanded leaves. Mean individual leaf area was then used to interpolate total seedling leaf area through time with weekly leaf counts. Root mass was collected by removing the roots system and passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots of seedlings in containers were not considered pot bound, as clusters of roots along the soil-container interface were not observed. Roots from the free seedlings were collected by excavating each 1 m^2^ subplot to the hard layer and keeping only roots within the subplot. For each seedling, a sub-sample of washed fine roots was analyzed for root length using WhinoRhizo software (Regent Instruments Inc., Quebec, QC, Canada). Specific root length (SRL) is reported as the root length divided by the dry mass of each sub-sample (m g^-1^). Fine root length density (FRLD) for seedlings in containers is reported as the total fine root length divided by the volume of each container (m dm^-3^).

## Photosynthetic parameters
Leaf gas exchange measurements were performed fortnightly at saturating light (A~sat~) and saturating light and [CO~2~] (A~max~) on new fully expanded leaves. Measurements were initiated only after sufficient new leaf growth occurred (March 05^th^, 2013), approximately 6 weeks following planting, and continued until the biomass harvest. Leaf level gas exchange was measured with a standard leaf chamber (2 x 3 cm) equipped with blue-red light emitting diodes using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA). A~sat~ measurements were made at PPFD of 1800 $\mu$mol m^-1^ s^-1^ and [CO~2~] of 400 $\mu$l l^-1^ and A~max~ with [CO~2~] of 1600 $\mu$l l^-1^ and PPFD of 1800 $\mu$mol photons m^-1^ s^-1^. This choice of light level to achieve light saturation is consistent with other studies on *Eucalyptus* species [@kallarackal1997ecophysiological;@pinkard1998photosynthetic;@crous2013photosynthesis;@drake2014capacity]. These measurements were conducted during midday (10:00-14:00 h) with leaf temperature maintained at 25 &deg;C. After CO~2~ and water vapor flux values stabilized in the leaf chamber, net CO~2~ assimilation rate and stomatal conductance (g~s~) were logged 5 times and averaged for both A~sat~ and A~max~.

@@

## Leaf water potential
Predawn ($\Psi$~pd~) and midday ($\Psi$~l~) leaf water potentials were measured for each seedling using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA) on fully expanded leaves during the same time period as AC~i~ and R. Leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. $\Psi$~pd~ was measured before sunrise and $\Psi$~l~ at midday 13:00-14:30 h. These measurements were used as a measure of static water stress on the seedlings [@sellin1999does] and to ensure that the bulk soil water availability was high enough for plants to avoid water stress as they became larger and roots filled the soil volume.

## Leaf, root and soil chemistry
Leaves used in each gas exchange measurements and subsamples of harvested roots were dried to a constant mass and milled for analysis of N content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were sieved to remove organic material, air dried and milled for analysis of N. Nitrogen concentrations of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer with thermal conductivity and mass spectromic detection (of N~2~ and CO~2~). The percentage of N in the sample was calculated by comparison with certified standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the VPDP scale.

@@

## Seedling growth model
We developed a simple seedling growth model that utilized leaf *A* rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~i~, stem~i~ and root~i~) and a starting leaf area (LA~i~) measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (P~i~) is then given by

@@

##Data analysis
Differences in experimental parameters with soil volume were analysed by mixed-effects models in R with individual containers and experimental blocks as random effects and soil volume treatment as a categorical fixed effect with seven levels. Tukey's post-hoc tests were performed in conjunction with ANOVA to determine which specific paired comparisons among soil volume treatments were different. A linear mixed effect model of A~max~ and leaf chemistry was performed using the 'nlme' package [@Pinheiro2015] in R. Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general]. Tests of allometric relationships between log-transformed biomass components were implemented using standardized major axis regression in the 'smatr' package in R [@warton2012smatr]. All tests of statistical significance were conducted at an alpha level of 0.05.

# Results

## Growth and morphology
Plant height, diameter and leaf area diverged between container volumes soon after start of the experiment (Figure `r allomfig`). First, seedling leaf area significantly diverged between soil volumes (P < 0.029) during the 5^th^ week of the experiment. Following this period both height (8^th^) week) and then diameter (9^th^ week) significantly deviated across soil volumes (P < 0.045 & 0.035, respectively). The large reductions in height gain and total leaf area in smaller compared to larger containers continued throughout the experiment. In this field study, colder temperatures and reductions in total PPFD per day (Figure `r airvarsfig`) likely slowed the growth of seedlings in the final weeks of the experiment (Figure `r allomfig`). Seedlings maintained diameter growth throughout the experiment, although marginal with smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (P < 0.001). Increases in both final stem diameter (P < 0.001) and cumulative leaf area (both P < 0.001) were found with increasing soil volume and these differences were driven mainly by the largest container and the free seedling treatments.
\
```{r allometry, fig.width=6, fig.height=7.8, echo=FALSE,  dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/allometryplotting.R")
```  
\
**Figure 2.1**. Soil volume treatment means &pm; standard error of height growth (a), diameter growth (b), and interpolated seedling leaf area (c) measured weekly of *Eucalyptus tereticornis* seedlings across the experiment duration in 2013.
\
```{r airvars, fig.width=6, fig.height=7.8, echo=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/airvars_plotting.R")
```  
\
**Figure 2.2**. Daily maximum and minimum temperature (a), total daily PPFD (b), and daily maximum vapour pressure deficit (c) across the experiment duration in 2013.

@@

## Leaf and root chemistry
Leaf N~f~ was significantly higher in free seedlings and the largest container volume compared to the smaller container volumes at the onset of gas exchange measurements (6th week, P < 0.001). Throughout the remainder of the experiment the smallest container volume had a significant reduction in leaf N~f~ compared to other soil volumes, while free seedlings maintained the highest leaf N~f~ (Table `r pvetab1`, P < 0.001). Leaf starch content in the smallest container was ca. double  that of free seedlings (P = 0.039), while leaf soluble sugars did not differ across treatments throughout the experiment (Table `r pvetab1`). Differences in leaf starch between the free seedling and the smallest container were evident during the first gas exchange campaign (P = 0.001). Root N was higher in free seedlings compared to seedlings in containers but only for some of the container size treatments (Table `r pvetab2`).

## Gas exchange and photosynthetic parameters
At the first measurement campaign, both A~sat~ and A~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both P < 0.001). Across all measurement campaigns mean A~sat~ (Figure `r asatfig`) and A~max~ (Table `r pvetab3`) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationships between photosynthetic capacity, leaf starch and leaf N on a mass basis was marginally significant (P = 0.058) but A~max~ on a mass basis was highly correlated to both leaf N content and leaf starch (both P < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationship of A~max~ to either leaf N content or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments A~max~ was higher when leaf N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, A~max~ was positively correlated with final harvest biomass across all seedlings (P < 0.001).
\
```{r asat, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE, message=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/A_stats_plotting.R")
```  
\
**Figure 2.4**. Soil volume treatment means &pm; standard error, across all measurement campaigns (n = 6), of light saturated rates of photosynthesis at 25&deg;C. Different letters represent significant differences between treatments.
\
```{r A_tncnitro, fig.width=8, fig.height=6, echo=FALSE,warning=FALSE, message=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/Amax_TNC_N_plotting2.R")
```  
\
**Figure 2.5**. Photosynthetic capacity, on a leaf mass basis, as a function of accumulation of leaf starch (a) and leaf nitrogen content without TNC (b).  Colors represent bins levels (n = 5) of both leaf starch and nitrogen grouped from low to high.  Lines represents predictions, for each bin level, from the linear mixed effects model equation of A~mass~ as a function of starch and nitrogen. The marginal R^2^ (fixed effects only) was 0.37 and the conditional R^2^ (fixed and random effects) was 0.48 for the complete model.

@@

## Modelling seedling biomass
The default model M~free~, successfully optimized a LMF (21.6 %) which then allowed the model to predict mean harvest total biomass of free seedlings within 1.2 %. Using this optimized LMF, the total biomass of modeled seedlings for each soil volume treatment (M~pots~) were on average 23&pm;2.4 g C more than measured seedling biomass when compared against predicted total net leaf C gain (Figure `r massmodelfig`a). Thus, seedling C mass was overestimated by an average of 50&pm;8.7 % in modeled seedlings across the soil volume treatments (Figure `r massmodelfig`b). As a result, the observed reductions in leaf *A* with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the  container size treatments.  
\
```{r massmodel, fig.width=6, fig.height=8, echo=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/massmodel_optimize.R")
```  
\
\
\
**Figure 2.6**. Total carbon mass for harvested and modeled seedlings versus predicted total carbon gain after 120 days (a) and  reductions in final seedling carbon mass, both modeled and observed, as a function of the reduction in leaf photosynthesis across treatments (b). For (a) the dashed 1:1 identifies the difference between net total leaf carbon gain and gross seedling production. For (b) both seedling carbon mass and daily carbon assimilation were first scaled to the free seedling control.

@@

# Discussion
This study utilized a simple but novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling biomass production. We addressed questions regarding the coordination of *A~n~* and growth by complementing empirical results with a C mass balance model. We found that reductions in leaf *A~n~* across container sizes, when integrated across the 120 day experiment, were alone insufficient to account for observed reductions in total plant biomass production.

## Reductions in growth and physiology under sink limitation
Soon after seedlings became established both height and diameter growth were negatively affected by decreasing soil volume. This led to the large reductions in biomass in small containers, compared to freely rooted seedlings. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, consistent with the meta-analysis of Poorter et al. [-@poorter2012pot]. These growth reductions were expected, as the impedance of root growth can cause reductions in overall plant growth and activity [@mcconnaughay1991physical; @young1997mechanical]. It has been shown that roots subjected to environmental stress may send inhibitory signals to the shoots that affect g~s~, cell expansion, cell division and the rate of leaf appearance [@passioura2002soil]. Here, this was evident in a large divergence in leaf area between seedlings in containers and free seedlings through time, with the eventual cessation of new leaf growth in seedlings in small containers.

@@

## Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, outside of ontogenetic drift (Figure `r allocfig`a,b). This is a surprising result as shifts in allocation have been noted specifically for nutrient limitation [@mcconnaughay1999biomass, and references therein]. Surprisingly, a constant ratio of fine root mass to leaf mass was observed across all treatments suggesting a functional partitioning response to optimize resource gain did not occur.

@@

## Do reductions in photosynthesis explain reductions in seedling growth?
Our model used a simple approach to drive seedling growth with measured reductions in leaf *A~n~* , via soil volume effects, while treating C use efficiency, respiration and C allocation as fixed processes. Contrary to expectation, the model consistently overestimated seedling growth in containers when parameterized with an optimized LMF for free seedlings. Although reductions in A~max~ and biomass were strongly correlated among treatments, as hypothesized by Poorter et al. [-@poorter2012pot], we provide evidence that the negative effects of sink limitation on *A~n~* can not fully explain explain reduced seedling growth. These findings are important as this model reflects classical approaches in tree growth and production modelling that are driven by inputs of C assimilation and processes such as respiration are considered proportional to biomass [@le2001carbon] or growth rate [@tjoelker1999acclimation]. It is possible that the overestimation of growth was due to an initial overestimation of *A~n~*, however, the robust empirical based methods used to generate photosynthetic parameters (J~max~, V~cmax~, R and g~1~) make this unlikely. 
Instead, our results indicate a need to evaluate how oversimplified representations of processes other than *A~n~* affect models which distinguish the fate of assimilate C within a plant. Doing so will provide valuable input to future models as assimilate allocation is a key component in carbon-balance driven  plant growth models, yet C partitioning remains a key weakness  [@lacointe2000carbon]. To address this issue, we utilized the flexibility of this model to test plausible fates of the pool of simulated non-biomass C unaccounted for with observed mass balance. Similar to Lohier et al. [-@lohier2014explaining] we manipulated processes contributing to modeled seedling C mass balance, including changes to leaf C allocation or non-leaf tissue respiration, to quantitatively test their respective influences on model predictions.

@@

## Conclusions
With a novel field-based design we detected a massive effect of container volume on seedling growth, not between containers but with naturally planted seedlings. This is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants has seldom been studied. Although biomass partitioning was conserved, our empirical and model results suggest that the amount of photosynthate allocated to non-biomass pools such as TNC or respiration were likely altered by sink inhibition. The debate over how rates of photosynthesis affect plant growth or to what degree these rates are instead controlled by growth has existed for decades [@sweet1966role]. Our combined empirical and modeling approach shows that when non-photosynthesis parameters were kept constant changes in *A~n~* were not able to fully to predict changes in growth, an important distinction often missed in studies that manipulate source/sink activity. 
K&#246;rner [-@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control *A~n~* through the demand for C. Although this may be true, our results infer that quantifying the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

# Supporting Information {.unnumbered}
**Table 2.S1**.Seedling growth model default parameters.

>>>>

---
title: |
  | Chapter 2 
  | Below-ground sink limitation alters growth and carbon balance of Eucalyptus seedlings
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle_thesis.docx
  pdf_document: default
  html_document:
    number_sections: yes

@@

<!-- \ -->
<!-- *Corresponding author*:  -->
<!-- Courtney Campany -->
<!-- E: courtneycampany@gmail.com -->

@@

#2.1  Abstract

@@

<!-- ## Keywords {.unnumbered} -->
<!-- photosynthesis, growth, sink regulation, carbon allocation, soil volume -->

#2.2  Introduction

@@

#2.3  Materials and Methods

##2.3.1 Experimental design
This experiment was located at the Hawkesbury Forest Experiment site in Richmond, NSW, Australia. Plots were located in an open cover paddock that was converted from native pasture grasses. Top soils at this site are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg^-1^ total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity. At this site a soil hard layer exists at ~1.0 m with a transition to heavy clay soils. The climate for the region is classified as sub-humid temperate.

@@

##2.3.2 Growth and morphology metrics
Seedlings were planted in summer (January 21^st^ 2013) and stem height, diameter at 15 cm and leaf count were measured weekly thereafter. Once the growth rate of individual plants had significantly declined a full biomass harvest was completed and the experiment ended (May 21^st^ 2013). Dry mass of leaves, stems, roots and total leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) were measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing total measured leaf area by total leaf count of only fully expanded leaves. Mean individual leaf area was then used to interpolate total seedling leaf area through time with weekly leaf counts. Root mass was collected by removing the roots system and passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots of seedlings in containers were not considered pot bound, as clusters of roots along the soil-container interface were not observed. Roots from the free seedlings were collected by excavating each 1 m^2^ subplot to the hard layer and keeping only roots within the subplot. For each seedling, a sub-sample of washed fine roots was analyzed for root length using WhinoRhizo software (Regent Instruments Inc., Quebec, QC, Canada). Specific root length (SRL) is reported as the root length divided by the dry mass of each sub-sample (m g^-1^). Fine root length density (FRLD) for seedlings in containers is reported as the total fine root length divided by the volume of each container (m dm^-3^).

##2.3.3 Photosynthetic parameters
Leaf gas exchange measurements were performed fortnightly at saturating light (A~sat~) and saturating light and [CO~2~] (A~max~) on new fully expanded leaves. Measurements were initiated only after sufficient new leaf growth occurred (March 05^th^, 2013), approximately 6 weeks following planting, and continued until the biomass harvest. Leaf level gas exchange was measured with a standard leaf chamber (2 x 3 cm) equipped with blue-red light emitting diodes using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA). A~sat~ measurements were made at PPFD of 1800 $\mu$mol m^-1^ s^-1^ and [CO~2~] of 400 $\mu$l l^-1^ and A~max~ with [CO~2~] of 1600 $\mu$l l^-1^ and PPFD of 1800 $\mu$mol photons m^-1^ s^-1^. This choice of light level to achieve light saturation is consistent with other studies on *Eucalyptus* species [@kallarackal1997ecophysiological;@pinkard1998photosynthetic;@crous2013photosynthesis;@drake2014capacity]. These measurements were conducted during midday (10:00-14:00 h) with leaf temperature maintained at 25 &deg;C. After CO~2~ and water vapor flux values stabilized in the leaf chamber, net CO~2~ assimilation rate and stomatal conductance (g~s~) were logged 5 times and averaged for both A~sat~ and A~max~.

@@

##2.3.4 Leaf water potential
Predawn ($\Psi$~pd~) and midday ($\Psi$~l~) leaf water potentials were measured for each seedling using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA) on fully expanded leaves during the same time period as AC~i~ and R. Leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. $\Psi$~pd~ was measured before sunrise and $\Psi$~l~ at midday 13:00-14:30 h. These measurements were used as a measure of static water stress on the seedlings [@sellin1999does] and to ensure that the bulk soil water availability was high enough for plants to avoid water stress as they became larger and roots filled the soil volume.

##2.3.5 Leaf, root and soil chemistry
Leaves used in each gas exchange measurements and subsamples of harvested roots were dried to a constant mass and milled for analysis of N content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were sieved to remove organic material, air dried and milled for analysis of N. Nitrogen concentrations of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer with thermal conductivity and mass spectromic detection (of N~2~ and CO~2~). The percentage of N in the sample was calculated by comparison with certified standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the VPDP scale.

@@

##2.3.6 Seedling growth model
We developed a simple seedling growth model that utilized leaf *A* rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~i~, stem~i~ and root~i~) and a starting leaf area (LA~i~) measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (P~i~) is then given by

@@

##2.3.7 Data analysis
Differences in experimental parameters with soil volume were analysed by mixed-effects models in R with individual containers and experimental blocks as random effects and soil volume treatment as a categorical fixed effect with seven levels. Tukey's post-hoc tests were performed in conjunction with ANOVA to determine which specific paired comparisons among soil volume treatments were different. A linear mixed effect model of A~max~ and leaf chemistry was performed using the 'nlme' package [@Pinheiro2015] in R. Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general]. Tests of allometric relationships between log-transformed biomass components were implemented using standardized major axis regression in the 'smatr' package in R [@warton2012smatr]. All tests of statistical significance were conducted at an alpha level of 0.05.

#2.4 Results

##2.4.1 Growth and morphology
Plant height, diameter and leaf area diverged between container volumes soon after start of the experiment (Figure `r allomfig`). First, seedling leaf area significantly diverged between soil volumes (P < 0.029) during the 5^th^ week of the experiment. Following this period both height (8^th^) week) and then diameter (9^th^ week) significantly deviated across soil volumes (P < 0.045 & 0.035, respectively). The large reductions in height gain and total leaf area in smaller compared to larger containers continued throughout the experiment. In this field study, colder temperatures and reductions in total PPFD per day (Figure `r airvarsfig`) likely slowed the growth of seedlings in the final weeks of the experiment (Figure `r allomfig`). Seedlings maintained diameter growth throughout the experiment, although marginal with smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (P < 0.001). Increases in both final stem diameter (P < 0.001) and cumulative leaf area (both P < 0.001) were found with increasing soil volume and these differences were driven mainly by the largest container and the free seedling treatments.
\
```{r allometry, fig.width=6, fig.height=7.8, echo=FALSE,  dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/allometryplotting.R")
```  
\
**Figure 2.1**. Soil volume treatment means &pm; standard error of height growth (a), diameter growth (b), and interpolated seedling leaf area (c) measured weekly of *Eucalyptus tereticornis* seedlings across the experiment duration in 2013.
\
```{r airvars, fig.width=6, fig.height=7.8, echo=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/airvars_plotting.R")
```  
\
**Figure 2.2**. Daily maximum and minimum temperature (a), total daily PPFD (b), and daily maximum vapour pressure deficit (c) across the experiment duration in 2013.

@@

##2.4.2 Leaf and root chemistry
Leaf N~f~ was significantly higher in free seedlings and the largest container volume compared to the smaller container volumes at the onset of gas exchange measurements (6th week, P < 0.001). Throughout the remainder of the experiment the smallest container volume had a significant reduction in leaf N~f~ compared to other soil volumes, while free seedlings maintained the highest leaf N~f~ (Table `r pvetab1`, P < 0.001). Leaf starch content in the smallest container was ca. double  that of free seedlings (P = 0.039), while leaf soluble sugars did not differ across treatments throughout the experiment (Table `r pvetab1`). Differences in leaf starch between the free seedling and the smallest container were evident during the first gas exchange campaign (P = 0.001). Root N was higher in free seedlings compared to seedlings in containers but only for some of the container size treatments (Table `r pvetab2`).

##2.4.3 Gas exchange and photosynthetic parameters
At the first measurement campaign, both A~sat~ and A~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both P < 0.001). Across all measurement campaigns mean A~sat~ (Figure `r asatfig`) and A~max~ (Table `r pvetab3`) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationships between photosynthetic capacity, leaf starch and leaf N on a mass basis was marginally significant (P = 0.058) but A~max~ on a mass basis was highly correlated to both leaf N content and leaf starch (both P < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationship of A~max~ to either leaf N content or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments A~max~ was higher when leaf N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, A~max~ was positively correlated with final harvest biomass across all seedlings (P < 0.001).
\
```{r asat, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE, message=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/A_stats_plotting.R")
```  
\
**Figure 2.4**. Soil volume treatment means &pm; standard error, across all measurement campaigns (n = 6), of light saturated rates of photosynthesis at 25&deg;C. Different letters represent significant differences between treatments.
\
```{r A_tncnitro, fig.width=8, fig.height=6, echo=FALSE,warning=FALSE, message=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/Amax_TNC_N_plotting2.R")
```  
\
**Figure 2.5**. Photosynthetic capacity, on a leaf mass basis, as a function of accumulation of leaf starch (a) and leaf nitrogen content without TNC (b).  Colors represent bins levels (n = 5) of both leaf starch and nitrogen grouped from low to high.  Lines represents predictions, for each bin level, from the linear mixed effects model equation of A~mass~ as a function of starch and nitrogen. The marginal R^2^ (fixed effects only) was 0.37 and the conditional R^2^ (fixed and random effects) was 0.48 for the complete model.

@@

##2.4.4 Modelling seedling biomass
The default model M~free~, successfully optimized a LMF (21.6 %) which then allowed the model to predict mean harvest total biomass of free seedlings within 1.2 %. Using this optimized LMF, the total biomass of modeled seedlings for each soil volume treatment (M~pots~) were on average 23&pm;2.4 g C more than measured seedling biomass when compared against predicted total net leaf C gain (Figure `r massmodelfig`a). Thus, seedling C mass was overestimated by an average of 50&pm;8.7 % in modeled seedlings across the soil volume treatments (Figure `r massmodelfig`b). As a result, the observed reductions in leaf *A* with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the  container size treatments.  
\
```{r massmodel, fig.width=6, fig.height=8, echo=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/massmodel_optimize.R")
```  
\
\
\
**Figure 2.6**. Total carbon mass for harvested and modeled seedlings versus predicted total carbon gain after 120 days (a) and  reductions in final seedling carbon mass, both modeled and observed, as a function of the reduction in leaf photosynthesis across treatments (b). For (a) the dashed 1:1 identifies the difference between net total leaf carbon gain and gross seedling production. For (b) both seedling carbon mass and daily carbon assimilation were first scaled to the free seedling control.

@@

#2.5  Discussion
This study utilized a simple but novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling biomass production. We addressed questions regarding the coordination of *A~n~* and growth by complementing empirical results with a C mass balance model. We found that reductions in leaf *A~n~* across container sizes, when integrated across the 120 day experiment, were alone insufficient to account for observed reductions in total plant biomass production.

##2.5.1 Reductions in growth and physiology under sink limitation
Soon after seedlings became established both height and diameter growth were negatively affected by decreasing soil volume. This led to the large reductions in biomass in small containers, compared to freely rooted seedlings. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, consistent with the meta-analysis of Poorter et al. [-@poorter2012pot]. These growth reductions were expected, as the impedance of root growth can cause reductions in overall plant growth and activity [@mcconnaughay1991physical; @young1997mechanical]. It has been shown that roots subjected to environmental stress may send inhibitory signals to the shoots that affect g~s~, cell expansion, cell division and the rate of leaf appearance [@passioura2002soil]. Here, this was evident in a large divergence in leaf area between seedlings in containers and free seedlings through time, with the eventual cessation of new leaf growth in seedlings in small containers.

@@

##2.5.2 Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, outside of ontogenetic drift (Figure `r allocfig`a,b). This is a surprising result as shifts in allocation have been noted specifically for nutrient limitation [@mcconnaughay1999biomass, and references therein]. Surprisingly, a constant ratio of fine root mass to leaf mass was observed across all treatments suggesting a functional partitioning response to optimize resource gain did not occur.

@@

##2.5.3 Do reductions in photosynthesis explain reductions in seedling growth?
Our model used a simple approach to drive seedling growth with measured reductions in leaf *A~n~* , via soil volume effects, while treating C use efficiency, respiration and C allocation as fixed processes. Contrary to expectation, the model consistently overestimated seedling growth in containers when parameterized with an optimized LMF for free seedlings. Although reductions in A~max~ and biomass were strongly correlated among treatments, as hypothesized by Poorter et al. [-@poorter2012pot], we provide evidence that the negative effects of sink limitation on *A~n~* can not fully explain explain reduced seedling growth. These findings are important as this model reflects classical approaches in tree growth and production modelling that are driven by inputs of C assimilation and processes such as respiration are considered proportional to biomass [@le2001carbon] or growth rate [@tjoelker1999acclimation]. It is possible that the overestimation of growth was due to an initial overestimation of *A~n~*, however, the robust empirical based methods used to generate photosynthetic parameters (J~max~, V~cmax~, R and g~1~) make this unlikely. 
Instead, our results indicate a need to evaluate how oversimplified representations of processes other than *A~n~* affect models which distinguish the fate of assimilate C within a plant. Doing so will provide valuable input to future models as assimilate allocation is a key component in carbon-balance driven  plant growth models, yet C partitioning remains a key weakness  [@lacointe2000carbon]. To address this issue, we utilized the flexibility of this model to test plausible fates of the pool of simulated non-biomass C unaccounted for with observed mass balance. Similar to Lohier et al. [-@lohier2014explaining] we manipulated processes contributing to modeled seedling C mass balance, including changes to leaf C allocation or non-leaf tissue respiration, to quantitatively test their respective influences on model predictions.

@@

##2.5.4 Conclusions
With a novel field-based design we detected a massive effect of container volume on seedling growth, not between containers but with naturally planted seedlings. This is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants has seldom been studied. Although biomass partitioning was conserved, our empirical and model results suggest that the amount of photosynthate allocated to non-biomass pools such as TNC or respiration were likely altered by sink inhibition. The debate over how rates of photosynthesis affect plant growth or to what degree these rates are instead controlled by growth has existed for decades [@sweet1966role]. Our combined empirical and modeling approach shows that when non-photosynthesis parameters were kept constant changes in *A~n~* were not able to fully to predict changes in growth, an important distinction often missed in studies that manipulate source/sink activity. 
K&#246;rner [-@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control *A~n~* through the demand for C. Although this may be true, our results infer that quantifying the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

#2.6  Supporting Information
**Table 2.S1**.Seedling growth model default parameters.
